### 2014-08-09 Docs/flowchart/flowchart.tex
more work on what we are actually solving

Here we outline the algorithm currently implemented in the code.  The
latest published reference for MAESTRO is the multilevel
paper \cite{multilevel}.  In this description, we make frequent
reference to paper~I \cite{lowMach}, paper~II \cite{lowMach2},
paper~III \cite{lowMach3}, and paper~IV \cite{lowMach4}.

@@

@@

Additionally, we carry species around which can react.  The creation and destruction
of the species is described by their create rate, $\omegadot_k$, and the species
are defined by their mass fractions, $X_k \equiv \rho_k / \rho$, giving
\begin{equation}
\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub) = \rho \omegadot_k
\end{equation}
and
\begin{equation}
\sum_k X_k = 1
\end{equation}

@@

@@

The compressible momentum equation (written in terms of velocity is):
\begin{equation}
\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla p = -\rho |g| \er
\end{equation}
Subtracting off the base state, and defining the perturbational
pressure (sometimes called the dynamic pressure) as $\pi = p -                                    
p_0$, and perturbational density as $\rho' = \rho - \rho_0$, we have:
\begin{equation}
\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla \pi = -\rho' |g| \er
\end{equation}
or
\begin{equation}
\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub + \frac{1}{\rho} \nabla \pi =
   -\frac{\rho^\prime}{\rho} |g| \er
\end{equation}
This is the form of the momentum equation that we solved in papers I--III and in the multilevel
paper.

Several authors \cite{KP:2012,VLBWZ:2013} explored the idea of energy
conservation in a low Mach number system and found that an additional term (which can
look like a buoyancy) is needed in the low Mach number formulation, yielding:
\begin{equation}
\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub +
   \frac{\beta_0}{\rho} \nabla \left (\frac{p^\prime}{\beta_0} \right ) =
   -\frac{\rho^\prime}{\rho} |g| \er
\end{equation}

We decompose the full velocity field into a base
state velocity, $w_0$, that governs the base state dynamics, and a
local velocity, $\Ubt$, that governs the local dynamics, i.e.,
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$.
The velocity evolution equations are then
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0                                                                                                                                        
  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
- \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting (\ref{eq:w0 divergence})
into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

%-----------------------------------------------------------------------------
\subsection{$\Gamma_1$ Variation Changes}

>>>>

The equation set and solution procedure used by \maestro\ has evolved
over time.  In this chapter, we outline the algorithm currently
implemented in the code.  The latest published reference for \maestro
is the multilevel paper \cite{multilevel}.  In this description, we
make frequent reference to paper~I \cite{lowMach},
paper~II \cite{lowMach2}, paper~III \cite{lowMach3}, and
paper~IV \cite{lowMach4}.

@@

The gravitational acceleration, $g$ is either constant or a point-mass
with a $1/r^2$ dependence (see \S \ref{sec:planarinvsqgravity}) for plane-parallel geometries, or a monopole
constructed by integrating the base state density for spherical
geometries.

@@

Additionally, we carry species around which can react.  The creation and destruction
of the species is described by their create rate, $\omegadot_k$, and the species
are defined by their mass fractions, $X_k \equiv \rho_k / \rho$, giving
\begin{equation}
\frac{\partial \rho X_k}{\partial t} + \nabla \cdot (\rho \Ub X_k) = \rho \omegadot_k
\end{equation}
and
\begin{equation}
\sum_k X_k = 1
\end{equation}

@@

In this constraint, $\gammabar$ is the lateral average of
$\Gamma_1 \equiv d\log p / d\log \rho |_s$.  Using the lateral average
here makes it possible to cast the constraint as a
divergence.  \cite{KP:2012} discuss the general case where we want to
keep the local variations of $\Gamma_1$ (and we explored this in paper
III).  We also look at this in \S~\ref{sec:flow:gamma1vary}.

@@

The compressible momentum equation (written in terms of velocity is):
\begin{equation}
\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla p = -\rho |g| \er
\end{equation}
Subtracting off the base state, and defining the perturbational
pressure (sometimes called the dynamic pressure) as $\pi = p - p_0$,
and perturbational density as $\rho' = \rho - \rho_0$, we have:
\begin{equation}
\rho \frac{\partial \Ub}{\partial t} + \rho \Ub \cdot \nabla \Ub + \nabla \pi = -\rho' |g| \er
\end{equation}
or
\begin{equation}
\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub + \frac{1}{\rho} \nabla \pi =
   -\frac{\rho^\prime}{\rho} |g| \er
\end{equation}
This is the form of the momentum equation that we solved in papers
I--IV and in the multilevel paper.

Several authors \cite{KP:2012,VLBWZ:2013} explored the idea of energy
conservation in a low Mach number system and found that an additional
term (which can look like a buoyancy) is needed in the low Mach number
formulation, yielding:
\begin{equation}
\frac{\partial \Ub}{\partial t} + \Ub \cdot \nabla \Ub +
   \frac{\beta_0}{\rho} \nabla \left (\frac{p^\prime}{\beta_0} \right ) =
   -\frac{\rho^\prime}{\rho} |g| \er
\end{equation}
This is the form that we enforce in \maestro, and the choice is controlled
by \runparam{use\_alt\_energy\_fix}.

We decompose the full velocity field into a base state velocity,
$w_0$, that governs the base state dynamics, and a local velocity,
$\Ubt$, that governs the local dynamics, i.e., 
\begin{equation}
\Ub = w_0(r,t)\eb_r + \Ubt(\xb,t).
\end{equation}
with
$\overline{(\Ubt\cdot\eb_r)} = 0$ and
$w_0 = \overline{(\Ub\cdot\eb_r)}$.
The velocity evolution equations are then \MarginPar{we need the $\beta_0$ here}
\begin{eqnarray}
\frac{\partial w_0}{\partial t} &=& -w_0\frac{\partial w_0}{\partial
  r} - \frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r},\label{eq:w0
  evolution}\\ \frac{\partial\Ubt}{\partial t} &=& -\Ub\cdot\nabla\Ubt
  - \left(\Ubt\cdot\eb_r\right)\frac{\partial w_0}{\partial r}\eb_r -
\frac{1}{\rho}\nabla\pi +
\frac{1}{\rho_0}\frac{\partial\pi_0}{\partial r}\eb_r -
\frac{\rho-\rho_0}{\rho}g\eb_r.\label{eq:utildeupd}
\end{eqnarray}
where $\pi_0$ is the base state component of the perturbational pressure.
By laterally averaging to equation (\ref{eq:U divergence}),
we obtain a divergence constraint for $w_0$:
\begin{equation}
\nabla\cdot(\beta_0 w_0 \eb_r) = \beta_0\left(\Sbar - \frac{1}{\gammabar p_0}\frac{\partial p_0}{\partial t}\right).\label{eq:w0 divergence}
\end{equation}
The divergence constraint for $\Ubt$ can be found by subtracting
(\ref{eq:w0 divergence}) into (\ref{eq:U divergence}), resulting in
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0\left(S-\Sbar\right).\label{eq:utilde divergence}
\end{equation}

@@

%-----------------------------------------------------------------------------
\subsection{\label{sec:flow:gamma1vary} $\Gamma_1$ Variation Changes}
